---
title: "Chess Tournament Table"
author: "Nabila Hossain"
date: "September 27, 2015"
output: html_document
---

__In this project, you're given a text file with chess tournament results where the information has some structure. Your job is to create an R Markdown file that generates a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players: Player's Name, Player's State, Total Number of Points, Player's Pre-Rating, and Average Pre Chess Rating of Opponents.__ 

####_The following are the steps that I took, into to creating the .CSV file with  Player's Name, Player's State, Total Number of Points, Player's Pre-Rating, and Average Pre Chess Rating of Opponents._<a name="a"/>   
  
* ####[Uploading the chess tournament .text, from my GitHub repository.](#1)  
* ####[Pull the information about the player's name and state, using regular expression.](#2)  
* ####[Extract the information about the player's total points and pre rating, using regular expression.](#3)  
* ####[Creating a table (table1) with the player's number, name, state, total points and pre rating.](#4)  
* ####[Extract the first row of information, from the text file and creating a table (rounds).](#5)  
* ####[Using the melt functionto combined the opponents' information from the seven rounds of into one column and creating a table (table2).](#6)  
* ####[Combined table1 and table2 using the `sqldf` package to obtain the opponents' pre-rating numbers (table3).](#7)
* ####[Found the opponents' average pre-ratings number for each player using the  aggregate function (table4).](#8) 
* ####[Created the final Chess Tournament Table using `sqldf` by joining table1 and table4 (Chess_Tournament).](#9)  
* ####[Creating a `.CSV` file in the local GitHub repository and the link to the online repository.](#10)  

<a name="1"/>__First I uploaded the Chess Tournament text file into my GitHub _Project 1_ repository. Using the package `RCurl` I uploaded the Chess Tournament from online.__
```{r}
library(RCurl)
chess <- getURL("https://raw.githubusercontent.com/nabilahossain/Class-IS607/master/Project%201/tournamentinfo.txt")
chess
```
```
[UP](#a)  


<a name="2"/>__Using the `stringr` package I obtained the player's name and state. Using `str_extraxt_all` and `str_replace_all` functions and regular expression I extracted and formatted the data from the .txt file.__
```{r}
library(stringr)
name <- unlist(str_extract_all(chess, "\\| [[:upper:]- ]{4,} \\|"))
name <- str_replace_all(name, pattern = "(\\| )|([[:space:]]{1,}\\|)", replacement = "")
head(name[25:30])

state <- unlist(str_extract_all(chess, "\\|[[:space:]]{1,}[[A-Z]]{2} \\|"))
state <- str_replace_all(state, pattern = "(\\|[[:space:]]{1,})|([[:space:]]{1,}\\|)", replacement = "")
head(state, 10)
```
[UP](#a)  

<a name="3"/>__Used `str_extraxt_all` and `str_replace_all` functions and regular expression to extract and format the information about the player's total number of points and the player's pre-rating.__
```{r}
total_points <- unlist(str_extract_all(chess, "\\|[[:digit:].[:digit:]]{3}[[:space:]]{1,}\\|"))
total_points <- str_replace_all(total_points, pattern = "(\\|)|([[:space:]]{1,}\\|)", replacement = "")
head(total_points, 10)

pre_rating <- unlist(str_extract_all(chess, "[:] [[:alnum:] ]{2,9}\\-\\>"))
pre_rating <- str_replace_all(pre_rating, pattern = "(\\: )|(\\s{1,}\\-\\>)|([O-Q]\\d{1,2})|(\\-\\>)", replacement = "")
pre_rating <- as.numeric(pre_rating)
head(pre_rating, 10)
```
[UP](#a)  

<a name="4"/>__I also extracted the player's number (to be used for combining tables), using the same functions as before. Then I created a table with player's number, name, state, total number of points and their pre-rating number. I titled the table as `table1` for future reference.__
```{r}
player_num <- unlist(str_extract_all(chess, "\\d{1,2}\\s\\|"))
player_num <- str_replace_all(player_num, pattern = "(\\s\\|)", replacement = "")
player_num <- as.numeric(player_num)
table1 <- data.frame(player_num = player_num, name = name, state = state, total_pts = total_points, pre_rating = pre_rating)
head(table1)
```
[UP](#a)  


<a name="5"/>__Here I also used `str_extraxt_all` and `str_replace_all` functions and regular expression. However here I used it to extract all the information found on the first line of the player's information in the Chess Tournment text file. I obtained the player's number, name, and total points and their opponent's number in round 1 through round 7. Then using `str_split_fixed` function I created a table called rounds with 10 columns of information, as you can see below.__
```{r}
r1 <- unlist(str_extract_all(chess, "\\d{1,2}\\s\\|(.+?)\\|[[:upper:]]{1}\\s{1,4}[[:digit:]]{0,3}\\|\n"))
r1 <- str_replace_all(r1, pattern = "(\\|[[:upper:]]{1} )", replacement = "\\|")
head(r1, 10)
r1 <- str_replace_all(r1, pattern = "([[:space:]]{2})|(\\|\n)", replacement = "")
rounds <- data.frame(str_split_fixed(r1, "\\|", 10))
names(rounds) <- c("num", "name", "total_pts", "round1", "round2", "round3", "round4", "round5", "round6", "round7")
tail(rounds, 4)
```
[UP](#a)  

<a name="6"/>__Then I used the `melt` function from the package `reshape2` to combine each player's opponents' information from seven rounds into one column. Then I created a table (table2) using the `subset` function to get rid of all the information where there was no opponent number. For example a player might have played only 5 games out of the seven. Therefore, only five of rounds would have any opponent information, while the other two have none.__
```{r}
library(reshape2)
r3 <- data.frame(rounds[c(1, 4:10)])
r3$num <- str_replace_all(r3$num, pattern = "\\s{1,}(\\d{1,2})", replacement = "\\1")
r4 <- melt(r3,id.vars="num", variable.name = "rounds", value.name = "opponent_number" )
tail(r4)
table2 <- subset(r4, opponent_number != " ")
table2$num <- as.numeric(table2$num)
tail(table2)
```
[UP](#a)  

 
<a name="7"/>__To obtain the opponent's pre-rating information, I used the package `sqldf` to join table2 with table1. I pulled the name and pre-rating from the table1, and the player's number, rounds they played and the opponent number from table2. By setting the opponent's number in table2 equaling to the player number in table1, I was able to obtain the opponent's pre-rating number and thier name as shown below. I set the new data frame as table3.__
```{r}
library(sqldf)
table3 <- sqldf("select t2.num as 'player_num', t1.name as 'opponent_name', t2.rounds, t2.opponent_number, t1.pre_rating as 'opponent_pre_rating'   from table2 t2
        left join table1 t1
         on t2.opponent_number =  t1.player_num
        order by t2.num asc")
head(table3)
```
[UP](#a)  

 
<a name="8"/>__Using `stats` package's `aggregate` function I found the average opponent's pre-rating. I set the subset with only the information about the player number and the average opponent's pre-rating as table4.__
```{r}
library(stats)
table4 <- aggregate(opponent_pre_rating ~ player_num, data = table3, FUN = 'mean')
head(table4)
```
[UP](#a)  

<a name="9"/>__Finally I create the table Chess_Tournament, with the following information: Player's Name, Player's State, Total Number of Points, Player's Pre-Rating, and Average Pre Chess Rating of Opponents. I use the `sqldf` function by joining table1 and table4. From table1 I obtain the player's name, player's state, total number of points, player's pre-rating, and from table4 I obtain average pre-rating of opponents. I also use the `format` function to round the decimals in average pre-rating of opponents column to 1 decimal place.__ 
```{r}
Chess_Tournament <- sqldf("select t1.name as 'Player_Name', t1.state as 'Player_State', t1.total_pts as 'Total_Points', t1.pre_rating as 'Player_Pre-Rating', t4.opponent_pre_rating as 'Opponents_Average_Pre_Rating'
      from table1 t1 left join table4 t4
         on t4.player_num =  t1.player_num")
Chess_Tournament$Opponents_Average_Pre_Rating <- format(round(Chess_Tournament$Opponents_Average_Pre_Rating, 1), nsmall = 1)
head(Chess_Tournament, 30)
```
[UP](#a)  


<a name="10"/>__I created a `.CSV` file in my local GitHub repository. You can view the CSV file in my online repository, when you right click on the link and open in new tab: [Chess Tournament CSV](https://github.com/nabilahossain/Class-IS607/blob/master/Project%201/Chess_Tournament.csv) or by copying and pasting the following link: <https://github.com/nabilahossain/Class-IS607/blob/master/Project%201/Chess_Tournament.csv>.__
```{r}
write.csv(Chess_Tournament, file = "C:/Users/Nabila/Documents/GitHub/Class-IS607/Project 1/Chess_Tournament.csv")
```
[UP](#a) 
