---
title: "Week 10 Assignment"
author: "Nabila Hossain"
date: "November 1, 2015"
output: html_document
---

####Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it to an R dataframe.

__I choose _The BOOK API_ from The New York Times web site: <http://developer.nytimes.com/docs>. I obtained (by registering) an API key from the web site. Then I used the instruction found in [Books API: Best Sellers](https://github.com/nabilahossain/Class-IS607/blob/master/Week%208%20Assignment/Week_8_Assignment_JSON.json) Link: <http://developer.nytimes.com/docs/books_api/Books_API_Best_Sellers> to read the file into R and create a dataframe.__

 
* [Use `jsonlite` package to obtain the list names.](#1a)  
* [Using the list name E-Book Fiction, I create a dataframe in R.](#1b)   
* [Use `httr` package to create the same dataframe.](#1c)  
* [Writing a .csv file in GitHub, and a table in R.](#1d)     
  
* <a name="1a"/>__Using the `jsonlite` package I read the JSON file that contains the list name (the name of the Times best-seller list). I did this step so that I can see the complete list of names found in the Book API.__

```{r, message=FALSE, warning=FALSE}
library(jsonlite)
res <- fromJSON("http://api.nytimes.com/svc/books/v3/lists/names?api-key=0ebfcb34c7334257471f7024c5ca0d48:8:73344009")
colnames(res$results)
list_name <- res$results$list_name
list_name

```

[UP](#a)  


* <a name="1b"/>__Using the results above I chose to create a dataframe, with the list of best-selling fiction e-books. Then using the pactage `jsonlite` I obtained the JSON API that contains the information about e-book fiction. With that information I created a datafrme with a subset of the original data (choose only the columns that I wanted).__

```{r}
res2 <- fromJSON("http://api.nytimes.com/svc/books/v3/lists/E-Book-Fiction?api-key=0ebfcb34c7334257471f7024c5ca0d48:8:73344009")
books_t0 <- data.frame(res2$results$books)
head(books_t0,3)
books_t1 <- data.frame(books_t0[c(1, 3, 11:13, 6:8)])
head(books_t1)
```

[UP](#a)  


* <a name="1c"/>__I wanted to use `httr` package, to obtain the Books API from online, as was shown in the textbook (for this class). Therefore, I used the `httr` package to obtain the Book API and used the package `jsonlite` to read the file and create a database. It look the same as only using the `jsonlite` package.__
```{r, message=FALSE, warning=FALSE}
library(httr)
link <- GET("http://api.nytimes.com/svc/books/v3/lists/E-Book-Fiction?api-key=0ebfcb34c7334257471f7024c5ca0d48:8:73344009", content_type_json())

#Using 'jsonlite' package:
jsonfile <- content(link, as = "text") 
res <- fromJSON(jsonfile)
books_t2 <- data.frame(res$results$books)
books_t3 <- data.frame(books_t2[c(1, 3, 11:13, 6:8)])
head(books_t3)
```

[UP](#a)  


* <a name="1d"/>__I created a CSV file with the subset of the books table 1. I also produced a table in R using the `knitr` package.__
```{r, message=FALSE, warning=FALSE}
write.csv(books_t1, file = "C:/Users/Nabila/Documents/GitHub/Class-IS607/Week 10 Assignment/books.csv")

library(knitr)
kable(books_t1, caption = "Table of the Top 15 New York Times Best Selling e-book fiction.", align = "c")
```


[UP](#a)  

